A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-05 latex/main.tex
Updates from Overleaf

To compare the performance of sc2ts and UShER+RIPPLES 
in terms of recombinant detection we ran RIPPLES on the Viridian UShER phylogeny,
subset to the 2,475,418 samples shared with the sc2ts ARG (STAR Methods).
Note that we are concerned with the direct output of sc2ts and UShER+RIPPLES, so QC filtering was not performed for either.
Setting the ``parsimony improvement'' parameter to $p=4$, equivalent to $k=4$ in sc2ts, 
RIPPLES reported 1174 recombination events.
Running 3SEQ on each of the candidate trios (STAR Methods) identified 601 of these as recombinant, giving a concordance rate of 51\%.
Aligning the events reported by both methods using the sets of descendant
samples (STAR Methods), 468 of the events are shared by both,
with the intersection comprising 40\% of the RIPPLES events
and 55\% of the sc2ts events (Figure~\ref{fig:ripples_sc2ts_events}). 
The 706 recombination events output by RIPPLES that do not 
correspond to a recombination event in the sc2ts ARG can largely
be explained by a combination of the choice of sites 
used for inference, and the strict time ordering imposed
by sc2ts (Document~\ref{sec:detailed_ripples}).
The 387 events detected by sc2ts and not by UShER+RIPPLES
would require a detailed dissection of the UShER tree
to explain, which is beyond the scope of this work.
However, the set does contain some well characterized events 
associated with the Pango X lineages.
Table~\ref{tab:ripples_x_lineages} shows the events detected 
by UShER+RIPPLES associated with samples assigned to Pango X
lineages. 
only 9 of the Pango X lineages are present and
with unequivocal recombinant origin for 6 
(contrast with 16 Type I events in Table~\ref{tab:pango_x_lineages}).
We therefore reran RIPPLES with $p=3$, resulting in 4111 events overall,
and increasing the intersection of events with sc2ts to 595 (70\%;
Figure~\ref{fig:ripples_sc2ts_events}).
At this more lenient threshold, 23 of the 44 Pango X lineages
are present in the dataset, with unequivocal origins for 10 lineages, and
7 lineages associated with two competing events 
(Document~\ref{sec:detailed_ripples}).
XB and XAJ are among the lineages detected by UShER+RIPPLES,
strengthening evidence that these are false negatives 
for sc2ts (Table~\ref{tab:pango_x_lineages}).

>>>>

To compare the performance of sc2ts and UShER+RIPPLES 
in terms of recombinant detection we ran RIPPLES on the Viridian UShER phylogeny,
subset to the 2,475,418 samples shared with the sc2ts ARG (STAR Methods).
Note that we are concerned with the direct output of sc2ts and UShER+RIPPLES, so QC filtering was not performed for either.
Setting the ``parsimony improvement'' parameter to $p=4$, equivalent to $k=4$ in sc2ts, 
RIPPLES reported 1174 recombination events.
Running 3SEQ on each of the candidate trios (STAR Methods) identified 601 of these as recombinant, giving a concordance rate of 51\%.
Aligning the events reported by both methods using the sets of descendant
samples (STAR Methods), 468 of the events are shared by both,
with the intersection comprising 40\% of the RIPPLES events
and 55\% of the sc2ts events (Figure~\ref{fig:ripples_sc2ts_events}). 
The 706 recombination events output by RIPPLES that do not 
correspond to a recombination event in the sc2ts ARG can largely
be explained by a combination of the choice of sites 
used for inference, and the strict time ordering imposed
by sc2ts (Document~\ref{sec:detailed_ripples}).
The 387 events detected by sc2ts and not by UShER+RIPPLES
would require a detailed dissection of the UShER tree
to explain, which is beyond the scope of this work.
However, the set does contain some well characterized events 
associated with the Pango X lineages.
Table~\ref{tab:ripples_x_lineages} shows the events detected 
by UShER+RIPPLES associated with samples assigned to Pango X
lineages, showing that at $p=4$
only 9 of the Pango X lineages are present and
with unequivocal recombinant origin for 6 
(contrast with 16 Type I events in Table~\ref{tab:pango_x_lineages}).
We therefore reran RIPPLES with $p=3$, resulting in 4111 events overall,
and increasing the intersection of events with sc2ts to 595 (70\%;
Figure~\ref{fig:ripples_sc2ts_events}).
At this more lenient threshold, 23 of the 44 Pango X lineages
are present in the dataset, with unequivocal origins for 10 lineages, and
7 lineages associated with two competing events 
(Document~\ref{sec:detailed_ripples}).
XB and XAJ are among the lineages detected by UShER+RIPPLES,
strengthening evidence that these are false negatives 
for sc2ts (Table~\ref{tab:pango_x_lineages}).
